---
title: "Bachelor Thesis"
author: "Oliver Tausendschön"
date: "2023-10-21"
output:
  pdf_document: default
  html_document: default
---
#Dataset:
```{r setup, include=FALSE}
df<-read.csv("NCD_RisC_Lancet_2017_BMI_age_standardised_country_adapted.csv")
```
```{r}
#combine male&female:
colnames(df)[8] <- "Obesity_Prevalence"
colnames(df) [23] <- "Overweight_Prevalence"
combined<-aggregate(cbind(Obesity_Prevalence, Overweight_Prevalence) ~ Country.Region.World + Year + ISO, df, mean)
combined["oao"]<-combined["Obesity_Prevalence"]+ combined["Overweight_Prevalence"]
countries<-c("EST", "FRA", "PER", "IND", "LVA", "COL", "IDN", "JPN", "LTU", "HUN", "SVN", "POL", "HRV", "RUS", "SVK", "CRI", "ROU", "CZE", "ZAF", "KOR", "CHE", "CHN", "NZL", "ISR", "AUT", "GBR", "LUX", "BGR", "ISL", "AUS", "CHL", "BRA", "MEX", "ITA", "IRL", "IRL", "SWE", "GRC", "DNK", "FIN", "BEL", "ESP", "NOR", "ARG", "PRT", "CYP", "MLT", "CAN", "DEU", "NLD", "TUR", "SAU", "USA")

result_oao<-subset(combined, ISO %in% countries)
```

#Direct medical costs
Costs of 2016:

```{r}
#install.packages("readxl")
#library(readxl)
#direct_medical_costs <- read_xlsx("Health_expenditure_52.xlsx")
#str(direct_medical_costs)
#updated_dataframe <- direct_medical_costs[-1, ]
#as.numeric(updated_dataframe$2019)
#medical_total<-sum(direct_medical_costs$2019[-1,])

direct_medical_costs<-read.csv("Health_expenditure_52_WHO.csv", dec = ".")
direct_medical_costs<-direct_medical_costs[-1,]
direct_medical_costs$X2019 <- as.numeric(gsub(",", "", direct_medical_costs$X2019))

medical_total<-sum((direct_medical_costs$X2019)) #costs in millions
medical_total*10^6 #costs
```

```{r}

```
#Project oao of 2019:
```{r}
# Fit a linear regression model
model <- lm(oao ~ Year + ISO, data = result_oao)
# Create a data frame with all combinations of ISO and Year for 2019
new_data <- expand.grid(Year = 2019, ISO = unique(result_oao$ISO))
# Make predictions for 2019
predicted_data <- data.frame(new_data, oao = predict(model, newdata = new_data))
predicted_data
```
Costs of 2016:
```{r}

```

#linear regression of oao / mean oao in 2020-2050 globally

```{r}
# Fit a linear regression model
model <- lm(oao ~ Year, data = result_oao)


# Predict the obesity prevalence for the years 2020 to 2050
years_to_predict <- data.frame(Year = 2020:2050)
predictions <- data.frame(
  Year = years_to_predict$Year,
  oao = predict(model, newdata = years_to_predict)
)


# Calculate the average obesity prevalence for the years 2020 to 2050
average_prevalence <- mean(predictions$oao)

average_prevalence

```
#linear regression of oao on all the countries and us it to then predict medical costs in 2020-2050:

```{r}
#prediction of every country:
projection <- data.frame()

for (country in unique(result_oao$Country.Region.World)){
  country_df<-result_oao[result_oao$Country.Region.World == country, ]
  
  model <- lm(oao ~ Year, data = country_df)
  years_to_predict <- data.frame(Year = 2020:2050)
  predictions <- data.frame(
    Year = years_to_predict$Year,
    oao = predict(model, newdata = years_to_predict)
  )
  
  # Calculate the average obesity prevalence for the years 2020 to 2050
  average_prevalence <- mean(predictions$oao)
  average_prevalence


  # Store the results in the 'results' data frame
  projection <- rbind(projection, data.frame(Country = country, Average_Prevalence = average_prevalence))
}

# Print the results
print(projection)
```


```{r}

#expand the model/regress it on direct medial costs so we can always plug in a number for obesity prevelance and then get oaf from that:
projection<-projection[order(projection$Average_Prevalence),]
print(projection)
projection["oafs"]=c(5.5834, 6.1217, 7.5428, 7.7690, 6.0866, 4.6819, 9.1332, 9.0287, 5.5180, 6.4475, 8.2022, 9.090, 6.906, 6.676, 11.337, 7.669, 4.878, 9.393, 10.065, 6.008, 8.536, 10.733, 6.449, 6.960, 6.022, 8.376, 8.691, 9.354, 9.743, 10.124, 9.714, 6.648, 7.063, 8.470, 6.246, 9.140, 6.195, 8.072, 9.033, 6.956, 7.534, 10.345, 8.613, 8.356, 9.910, 8.586, 8.934, 10.628, 7.920, 11.576, 13.531, 12.652)
projection

final<-lm(projection$Average_Prevalence~projection$oafs)
final
final$coefficients%*%c(1,13)
summary(final)

#should be correct:
final<-lm(projection$oafs~projection$Average_Prevalence)
final
final$coefficients%*%c(1,0.5)
summary(final)
#why is my coefficient different to the one in the paper?
plot(projection$oaf~projection$Average_Prevalence)
abline(final$coefficients)

```

#direct non-medical costs:
##travel cost:
```{r}
gas<-read.csv("e2373858-926a-4755-8350-78bb920f6ee9_Data.csv")
gas <- gas[, c(1,2,3,4,50,51,52,53,54,55,56,57,58, 59, 60, 61)]
gas$gas2016<-gas$X2016..YR2016.
#replace missing values by mean:
gas$gas2016[gas$gas2016 == ".."] <- NA
m <- mean(as.numeric(gas$gas2016), na.rm = TRUE)
gas$gas2016[is.na(gas$gas2016)] <- m
#reduce it to countries used above:
countries<-c("EST", "FRA", "PER", "IND", "LVA", "COL", "IDN", "JPN", "LTU", "HUN", "SVN", "POL", "HRV", "RUS", "SVK", "CRI", "ROU", "CZE", "ZAF", "KOR", "CHE", "CHN", "NZL", "ISR", "AUT", "GBR", "LUX", "BGR", "ISL", "AUS", "CHL", "BRA", "MEX", "ITA", "IRL", "IRL", "SWE", "GRC", "DNK", "FIN", "BEL", "ESP", "NOR", "ARG", "PRT", "CYP", "MLT", "CAN", "DEU", "NLD", "TUR", "SAU", "USA")

result_gas<-subset(gas, Country.Code %in% countries)
head(result_gas)
```

```{r}
population<-read.csv("population_WHO.csv")
population <- population[, c(1,2,3,4,50,51,52,53,54,55,56,57,58, 59, 60, 61)]

#reduce it to countries used above:
result_population<-subset(population, Country.Code %in% countries)
head(result_population)

```
```{r}
#parameter sourced from a scoping review of peer-reviewed studies!
#average outpatient visits per capita
outpatient<-5.42
#average inpatient visits per capita
inpatient<-0.10
```
calculate for 2016:
```{r}
#we first have to order all datasets so the used valus match the corresponding country:
result_gas<-result_gas[order(result_gas$Country.Code), ]
result_population<-result_population[order(result_population$Country.Code),]
result_oao_2016<-result_oao[result_oao$Year == "2016", ]
result_oao_2016<-result_oao_2016[order(result_population$Country.Code),]
result_oao_2016$population_with_obesity<-as.numeric(result_population$X2016..YR2016.)*as.numeric(result_oao_2016$oao)

#gas price per l * km * l/km:
result_gas$atc<-as.numeric(result_gas$gas2016)*5*0.06

outpatient_travel_cost<-result_gas$atc*outpatient*result_oao_2016$population_with_obesity
inpatient_travel_cost<-result_gas$atc*inpatient*result_oao_2016$population_with_obesity
outpatient_travel_cost
inpatient_travel_cost
travel_cost<-outpatient_travel_cost+inpatient_travel_cost
travel_cost
```

##Informal caregiver travel cost

```{r}
caregiver_travel_cost<-result_gas$atc*inpatient*result_oao_2016$population_with_obesity
total_travel_cost<-caregiver_travel_cost+travel_cost

```

#indirect costs:

##premature mortality
Caculate how many years oao reduces life expectancy/how many years of life. We can do a regression of oao population on obesity prevelance?
For simplicity, we now take 5years as a proxy for years lost due to premature oao-related deaths. Furthermore, we adjust GDP per capita for the full income value of a life year, which combines changes in national income and the value of changes in mortality or life expectancy by using 1.6 as a GDP multiplier for all countries.
```{r}
gdp<-read.csv("GDP_Worldbank.csv")
result_gdp<-subset(gdp, Country.Code %in% countries)
result_gdp<-result_gdp[order(result_gdp$Country.Code),]

head(result_gdp)
```


```{r}
cost_OAO_attributable_mortality<-(as.numeric(result_gdp$X2016..YR2016.)*1.6*5*as.numeric(result_oao_2016$population_with_obesity))
cost_OAO_attributable_mortality
```

##reduced productivity due to additional absenteeism:
For simplicity, we take 75% as a proxy for the percentage of working age population:

```{r}
income<-read.csv("income_Worldbank.csv")
result_income<-subset(income, Country.Code %in% countries)
result_income<-result_income[order(result_income$Country.Code),]
head(result_income)
result_income$daily_2016<-as.numeric(result_income$X2016..YR2016.)/365
#no income for malta!
```

```{r}
#parameter sourced from a scoping review of peer-reviewed studies!
additional_days_absent<-4
working_population<-0.75
average_daily_wage<-result_income$daily_2016

absenteeism_cost<-result_oao_2016$population_with_obesity*0.75*additional_days_absent*average_daily_wage

absenteeism_cost
```

##reduced productivity due to additional presenteeism:

45, 58, 62, 49, 53, 45
rate of reduced productivity? e.g. 2%?
```{r}
additional_presenteeism_rate<-0.12
annual_wage<-as.numeric(result_income$X2016..YR2016.)

presenteeism_cost<-result_oao_2016$population_with_obesity*0.75*additional_presenteeism_rate*annual_wage
presenteeism_cost

```
##total costs:

```{r}

total_costs<-presenteeism_cost + absenteeism_cost + cost_OAO_attributable_mortality + total_travel_cost
total_costs
sum(total_costs, na.rm = T)
```

