---
title: "Thesis_v0.9"
author: "Oliver Tausendschön"
date: "2023-11-06"
bibliography: references.bib
link-citations: yes
output:
  pdf_document: default
  html_document: default
---

```{r, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = FALSE)
```



# Appendix

# Dataset Overweight and Obesity:

```{r}
data_oao<-read.csv("NCD_RisC_Lancet_2017_BMI_age_standardised_country_adapted.csv")
```

In order to work with the data, we make some adjustments to the dataset:
underweight: bmi <18.5
under-normal weight: 18.5<bmi<20
normal weight: 20<bmi<25
overweight: 25<bmi<30
moderate obesity 30<bmi<35
obesity: bmi >=30
severe obesity bmi >= 35
```{r}
colnames(data_oao)[8]<-"obesity"
colnames(data_oao)[11]<-"severe_obesity"
colnames(data_oao)[14]<-"underweight"
colnames(data_oao)[17] <-"under-normal_weight"
colnames(data_oao)[20] <-"normal_weight"
colnames(data_oao)[23]<-"overweight"
colnames(data_oao)[26]<-"moderate_obesity"
data_oao<-data_oao[, -c(6,7,9,10,12,13,15,16,18,19,21,21,22,24,25,27,28,29,30,31,32,33,34)]
#create healthy weight
data_oao["healthy_weight"]<-data_oao$underweight + data_oao$`under-normal_weight` + data_oao$normal_weight
#create oao
data_oao["oao"]<-data_oao["obesity"]+ data_oao["overweight"]
data_oao_men<-data_oao[data_oao$Sex == "Men",]
data_oao_women<-data_oao[data_oao$Sex == "Women",]

```

```{r}
countries<-c("EST", "FRA", "PER", "IND", "LVA", "COL", "IDN", "JPN", "LTU", "HUN", "SVN", "POL", "HRV", "RUS", "SVK", "CRI", "ROU", "CZE", "ZAF", "KOR", "CHE", "CHN", "NZL", "ISR", "AUT", "GBR", "LUX", "BGR", "ISL", "AUS", "CHL", "BRA", "MEX", "ITA", "IRL", "SWE", "GRC", "DNK", "FIN", "BEL", "ESP", "NOR", "ARG", "PRT", "CYP", "MLT", "CAN", "DEU", "NLD", "TUR", "SAU", "USA")
oafs<-sort(c(5.5834, 6.1217, 7.5428, 7.7690, 6.0866, 4.6819, 9.1332, 9.0287, 5.5180, 6.4475, 8.2022, 9.090, 6.906, 6.676, 7.669, 4.878, 9.393, 10.065, 6.008, 8.536, 10.733, 6.449, 6.960, 6.022, 8.376, 8.691, 9.354, 9.743, 10.124, 9.714, 6.648, 7.063, 8.470, 6.246, 9.140, 6.195, 8.072, 9.033, 6.956, 7.534, 10.345, 8.613, 8.356, 9.910, 8.586, 8.934, 10.628, 7.920, 11.337, 11.576, 12.652, 13.531)/100)

Country <- c("Estonia","France","Peru","India","Latvia","Colombia","Indonesia","Japan","Lithuania", "Hungary", "Slovenia", "Poland", "Croatia", "Russian Federation", "Slovakia","Costa Rica","Romania","Czech Republic","South Africa","South Korea", "Switzerland","China","New Zealand","Israel","Austria","United Kingdom","Luxembourg","Bulgaria","Iceland","Australia","Chile","Brazil","Mexico","Italy","Ireland","Sweden","Greece","Denmark","Finland","Belgium","Spain","Norway","Argentina","Portugal","Cyprus","Malta","Canada","Germany","Netherlands","Turkey","Saudi Arabia","United States of America")
```

oao for boys and girls:

```{r}

data_oao_child<-read.csv("NCD_RisC_Lancet_2020_BMI_child_adolescent_country.csv")
colnames(data_oao_child)[12]<-"oao"
data_oao_child<-data_oao_child[, c(1,2,3,4,5,12)]
data_oao_child <- aggregate(oao ~ Country + Sex + Year, data = data_oao_child, FUN = mean) #we take the mean because we suppose that there are equally many indiviudals in each age-group
data_oao_child_girls<-data_oao_child[data_oao_child$Sex == "Girls", ]
data_oao_child_boys<-data_oao_child[data_oao_child$Sex == "Boys", ]
```

# oao prevelance projection: autoregression

## exogenous variables

```{r}
#add external proxies to develop ARIMAX (population change):

population_per<-read.csv("WPP2022_PopulationByAge5GroupSex_Percentage_Medium.csv") #turkey missing?
colnames(population_per)[13]<-"Year"
colnames(population_per)[4] <- "ISO"
colnames(population_per)[8] <- "Country"
population_per<-population_per[population_per$Year >= "1975" & population_per$Year<="2060", ]
population_per<-subset(population_per, ISO %in% unique(data_oao$ISO))
#create Dataframes for all, child, and adults, always using the aggregate to use agegroups.
ag_population_all<-aggregate(cbind(PopTotal, PopFemale, PopMale) ~ ISO + Year + Country + Location, data = population_per, sum)
ag_population_child<-population_per[population_per$AgeGrp %in% c("0-4", "5-9", "10-14", "15-19"), ]
ag_population_child <- aggregate(cbind(PopTotal, PopFemale, PopMale) ~ ISO + Year + Country + Location, data = ag_population_child, sum)
ag_population_adult <- population_per[!(population_per$AgeGrp %in% c("0-4", "5-9", "10-14", "15-19")), ]
ag_population_adult <- aggregate(cbind(PopTotal, PopFemale, PopMale) ~ ISO + Year + Country + Location, data = ag_population_adult, sum)


#add external proxies to develop ARIMAX (urbanization):
urban_per<-read.csv("AnnualUrbanPopMidYear-2050.csv",skip = 1, header = T) #only until 2050? We decided to predict oao prevelance without this factor as it only ranges to 2050 and not 2060.
urban_per$Location<-trimws(urban_per$Location)
urban_per<-subset(urban_per, Location %in% unique(data_oao$Country.Region.World))#no ISO --> A FEW COUNTRIES MISSING
urban_per<-urban_per[, -c(1,3)]

library(tidyr)
# Reshape the dataframe
urban_per <- pivot_longer(urban_per, cols = -Location, names_to = "Year", values_to = "UrbanPopulationPercentage")
urban_per$Year <- gsub("X", "", urban_per$Year)
urban_per <- urban_per[urban_per$Location != "Tokelau", ]
urban_per <- urban_per[urban_per$Location != "Nauru", ]
urban_per <- urban_per[urban_per$Location != "Singapore", ]
urban_per <- urban_per[urban_per$Location != "Bermuda", ]
urban_per <-urban_per[urban_per$Year!="2017",] #remove 2017 and 2018 to start with 2019
urban_per <-urban_per[urban_per$Year!="2018",]
library(dplyr)
urban_per$Year<-as.integer(urban_per$Year)

```




## autoregression of 161 countries using c(2,1,0) and using exception .notnull with  c(2,2,0)

This is the model used to predict prevalence. I still have to experiment with the model parameters to find out the optimal combination. But for now I decided with to stick with these as the others did not work...
So we fit an arimax(2,1,0) model and do this for men, women, girls and boys. We decided to leave out urbanization as a factor to enable us predicting until 2060. (data on urbanization only goes up to 2050)

### Loop for men:
```{r}
library(forecast)

forecast_df_men <- data.frame()
m_2020_2050_men <- data.frame()
for (country in unique(urban_per$Location)) {
  country_df <- data_oao_men[data_oao_men$Country.Region.World == country, ]
  ag_popu_df <- ag_population_adult[ag_population_adult$Location == country, ]
  ag_popu_df2<-ag_popu_df[ag_popu_df$Year>=2019, ]
  ag_popu_df<-ag_popu_df[ag_popu_df$Year<=2016, ]
  print(country)

  # Extract the time series and exogenous variables
  country_df$oao_logit <- log(country_df$oao / (1-country_df$oao))
  country_df$oao_log <- log(country_df$oao)

  ts_data <- ts(country_df$oao_logit, start = min(country_df$Year), frequency = 1)
  xreg_data <-ag_popu_df$PopMale
  xreg_data2<-ag_popu_df2$PopMale


  # Initialize arimax_model outside the tryCatch block
  arimax_model <- NULL

  tryCatch({
    # Try fitting ARIMAX model
    arimax_model <- Arima(ts_data, order = c(2, 1, 0), xreg = xreg_data)
  }, error = function(e) {
    # If an error occurs, print the error message
    print(paste(country, "Error1:", e))
  })

  if (!is.null(arimax_model)) {
    # Generate forecasts using the fitted model
    forecasts <- forecast(arimax_model, xreg = xreg_data2, h = 42)
    forecasts_original_scale <- exp(forecasts$mean) / (1 + exp(forecasts$mean))
    #forecasts_original_scale <- exp(forecasts$mean)

    # Create a data frame for forecasted values
    forecast_df_country <- data.frame(Country = rep(country, 42), Year = seq(2019, 2060), Forecast_men = as.numeric(forecasts_original_scale))
    forecast_df_men <- rbind(forecast_df_men, forecast_df_country)

    # Calculate the mean projection for 2020-2060
    country_mean <- mean(forecasts_original_scale[1:32])  # Adjust to use values for the years 2020-2050
    m_2020_2050_men <- rbind(m_2020_2050_men, data.frame(Country = country, MeanProjection = country_mean))
  }
 if (is.null(arimax_model)) {
  tryCatch({
    arimax_model <- Arima(ts_data, order = c(2, 2, 0), xreg = xreg_data)
    # Generate forecasts using the fitted model
    forecasts <- forecast(arimax_model, xreg = xreg_data2, h = 42)
    forecasts_original_scale <- exp(forecasts$mean) / (1 + exp(forecasts$mean))
    #forecasts_original_scale <- exp(forecasts$mean)


    # Create a data frame for forecasted values
    forecast_df_country <- data.frame(Country = rep(country, 42), Year = seq(2019, 2060), Forecast_men = as.numeric(forecasts_original_scale))
    forecast_df_men <- rbind(forecast_df_men, forecast_df_country)

    # Calculate the mean projection for 2020-2060
    country_mean <- mean(forecasts_original_scale[1:32])  # Adjust to use values for the years 2020-2050
    m_2020_2050_men <- rbind(m_2020_2050_men, data.frame(Country = country, MeanProjection = country_mean))
  }, error = function(e) {
    # Print an error message if the code block fails
    print(paste("Error2 in block:", e))
  })
  }
}


# Print the results
head(forecast_df_men)
print(m_2020_2050_men)
```



### Loop for Women:
```{r}
library(forecast)

forecast_df_women <- data.frame()
m_2020_2050_women <- data.frame()
for (country in unique(urban_per$Location)) {
  country_df <- data_oao_women[data_oao_women$Country.Region.World == country, ]
  urban_df <- urban_per[urban_per$Location == country, ]
  ag_popu_df <- ag_population_adult[ag_population_adult$Location == country, ]
  ag_popu_df2<-ag_popu_df[ag_popu_df$Year>=2019, ]
  ag_popu_df<-ag_popu_df[ag_popu_df$Year<=2016, ]

  
  print(country)

  # Extract the time series and exogenous variables
  country_df$oao_logit <- log(country_df$oao / (1-country_df$oao))
  ts_data <- ts(country_df$oao_logit, start = min(country_df$Year), frequency = 1)
  

  xreg_data <- ag_popu_df$PopFemale
  xreg_data2 <- ag_popu_df2$PopFemale


  # Initialize arimax_model outside the tryCatch block
  arimax_model <- NULL

  tryCatch({
    # Try fitting ARIMAX model
    arimax_model <- Arima(ts_data, order = c(2, 1, 0), xreg = xreg_data)
  }, error = function(e) {
    # If an error occurs, print the error message
    print(paste("Error1:", e))
  })

  if (!is.null(arimax_model)) {
    # Generate forecasts using the fitted model
    forecasts <- forecast(arimax_model, xreg = xreg_data2, h = 42)
    forecasts_original_scale <- exp(forecasts$mean) / (1 + exp(forecasts$mean))
    #forecasts_original_scale <- exp(forecasts$mean)

    # Create a data frame for forecasted values
    forecast_df_country <- data.frame(Country = rep(country, 42), Year = seq(2019, 2060), Forecast_women = as.numeric(forecasts_original_scale))
    forecast_df_women <- rbind(forecast_df_women, forecast_df_country)

    # Calculate the mean projection for 2020-2060
    country_mean <- mean(forecasts_original_scale[1:32])  # Adjust to use values for the years 2020-2050
    m_2020_2050_women <- rbind(m_2020_2050_women, data.frame(Country = country, MeanProjection = country_mean))
  }
 if (is.null(arimax_model)) {
  tryCatch({
    arimax_model <- Arima(ts_data, order = c(2, 2, 0), xreg = xreg_data)
    # Generate forecasts using the fitted model
    forecasts <- forecast(arimax_model, xreg = xreg_data2, h = 42)
    forecasts_original_scale <- exp(forecasts$mean) / (1 + exp(forecasts$mean))
    #forecasts_original_scale <- exp(forecasts$mean)


    # Create a data frame for forecasted values
    forecast_df_country <- data.frame(Country = rep(country, 42), Year = seq(2019, 2060), Forecast_women = as.numeric(forecasts_original_scale))
    forecast_df_women <- rbind(forecast_df_women, forecast_df_country)

    # Calculate the mean projection for 2020-2060
    country_mean <- mean(forecasts_original_scale[1:32])  # Adjust to use values for the years 2020-2050
    m_2020_2050_women <- rbind(m_2020_2050_women, data.frame(Country = country, MeanProjection = country_mean))
  }, error = function(e) {
    # Print an error message if the code block fails
    print(paste("Error2 in block:", e))
  })
  }
}

# Print the results
head(forecast_df_women)
print(m_2020_2050_women)
```

### Loop for boys:

```{r}
library(forecast)

forecast_df_boys <- data.frame()
m_2020_2050_boys <- data.frame()
for (country in unique(urban_per$Location)) {
  country_df <- data_oao_child_boys[data_oao_child_boys$Country == country, ]
  #urban_df <- urban_per[urban_per$Location == country, ]
  ag_popu_df <- ag_population_child[ag_population_child$Location == country, ]
  ag_popu_df2<-ag_popu_df[ag_popu_df$Year>=2019, ]
  ag_popu_df<-ag_popu_df[ag_popu_df$Year<=2016, ]

  print(country)

  # Extract the time series and exogenous variables
  country_df$oao_logit <- log(country_df$oao / (1-country_df$oao))
  country_df$oao_log <- log(country_df$oao)

  ts_data <- ts(country_df$oao_logit, start = min(country_df$Year), frequency = 1)
  #xreg_data <- cbind(urban_df$UrbanPopulationPercentage, ag_popu_df$PopMale)
  xreg_data<-ag_popu_df$PopMale
  xreg_data2<-ag_popu_df2$PopMale

  # Initialize arimax_model outside the tryCatch block
  arimax_model <- NULL

  tryCatch({
    # Try fitting ARIMAX model
    arimax_model <- Arima(ts_data,order = c(2, 1, 0), xreg = xreg_data)
  }, error = function(e) {
    # If an error occurs, print the error message
    print(paste("Error1:", e))
  })

  if (!is.null(arimax_model)) {
    # Generate forecasts using the fitted model
    forecasts <- forecast(arimax_model, xreg = xreg_data2, h = 42)
    forecasts_original_scale <- exp(forecasts$mean) / (1 + exp(forecasts$mean))
    #forecasts_original_scale <- exp(forecasts$mean)

    # Create a data frame for forecasted values
    forecast_df_country <- data.frame(Country = rep(country, 42), Year = seq(2019, 2060), Forecast_boys = as.numeric(forecasts_original_scale))
    forecast_df_boys <- rbind(forecast_df_boys, forecast_df_country)

    # Calculate the mean projection for 2020-2060
    country_mean <- mean(forecasts_original_scale[1:32])  # Adjust to use values for the years 2020-2050
    m_2020_2050_boys <- rbind(m_2020_2050_boys, data.frame(Country = country, MeanProjection = country_mean))
  }
 if (is.null(arimax_model)) {
  tryCatch({
    arimax_model <- Arima(ts_data, order = c(2, 2, 0), xreg = xreg_data)
    # Generate forecasts using the fitted model
    forecasts <- forecast(arimax_model, xreg = xreg_data2, h = 42)
    forecasts_original_scale <- exp(forecasts$mean) / (1 + exp(forecasts$mean))
    #forecasts_original_scale <- exp(forecasts$mean)


    # Create a data frame for forecasted values
    forecast_df_country <- data.frame(Country = rep(country, 42), Year = seq(2019, 2060), Forecast_boys = as.numeric(forecasts_original_scale))
    forecast_df_boys <- rbind(forecast_df_boys, forecast_df_country)

    # Calculate the mean projection for 2020-2060
    country_mean <- mean(forecasts_original_scale[1:32])  # Adjust to use values for the years 2020-2050
    m_2020_2050_boys <- rbind(m_2020_2050_boys, data.frame(Country = country, MeanProjection = country_mean))
  }, error = function(e) {
    # Print an error message if the code block fails
    print(paste("Error2 in block:", e))
  })
  }
}

# Print the results
head(forecast_df_boys)
print(m_2020_2050_boys)
```

### Loop for girls:
```{r}
library(forecast)

forecast_df_girls <- data.frame()
m_2020_2050_girls <- data.frame()
for (country in unique(urban_per$Location)) {
  country_df <- data_oao_child_girls[data_oao_child_girls$Country == country, ]
  urban_df <- urban_per[urban_per$Location == country, ]
  ag_popu_df <- ag_population_child[ag_population_child$Location == country, ]
  ag_popu_df2<-ag_popu_df[ag_popu_df$Year>=2019, ]
  ag_popu_df<-ag_popu_df[ag_popu_df$Year<=2016, ]

  print(country)

  # Extract the time series and exogenous variables
  country_df$oao_logit <- log(country_df$oao / (1-country_df$oao))
  country_df$oao_log <- log(country_df$oao)


  ts_data <- ts(country_df$oao_logit, start = min(country_df$Year), frequency = 1)
  xreg_data <- ag_popu_df$PopFemale
  xreg_data2<-ag_popu_df$PopFemale
  # Initialize arimax_model outside the tryCatch block
  arimax_model <- NULL

  tryCatch({
    # Try fitting ARIMAX model
    arimax_model <- Arima(ts_data, order = c(2, 1, 0), xreg = xreg_data)
  }, error = function(e) {
    # If an error occurs, print the error message
    print(paste("Error1:", e))
  })

  if (!is.null(arimax_model)) {
    # Generate forecasts using the fitted model
    forecasts <- forecast(arimax_model, xreg = xreg_data2, h = 42)
    forecasts_original_scale <- exp(forecasts$mean) / (1 + exp(forecasts$mean))
    #forecasts_original_scale <- exp(forecasts$mean)


    # Create a data frame for forecasted values
    forecast_df_country <- data.frame(Country = rep(country, 42), Year = seq(2019, 2060), Forecast_girls = as.numeric(forecasts_original_scale))
    forecast_df_girls <- rbind(forecast_df_girls, forecast_df_country)

    # Calculate the mean projection for 2020-2060
    country_mean <- mean(forecasts_original_scale[1:32])  # Adjust to use values for the years 2020-2050
    m_2020_2050_girls <- rbind(m_2020_2050_girls, data.frame(Country = country, MeanProjection = country_mean))
  }
 if (is.null(arimax_model)) {
  tryCatch({
    arimax_model <- Arima(ts_data, order = c(2, 2, 0), xreg = xreg_data)
    # Generate forecasts using the fitted model
    forecasts <- forecast(arimax_model, xreg = xreg_data2, h = 42)
    forecasts_original_scale <- exp(forecasts$mean) / (1 + exp(forecasts$mean))
    #forecasts_original_scale <- exp(forecasts$mean)


    # Create a data frame for forecasted values
    forecast_df_country <- data.frame(Country = rep(country, 42), Year = seq(2019, 2060), Forecast_girls = as.numeric(forecasts_original_scale))
    forecast_df_girls <- rbind(forecast_df_girls, forecast_df_country)

    # Calculate the mean projection for 2020-2060
    country_mean <- mean(forecasts_original_scale[1:32])  # Adjust to use values for the years 2020-2050
    m_2020_2050_girls <- rbind(m_2020_2050_girls, data.frame(Country = country, MeanProjection = country_mean))
  }, error = function(e) {
    # Print an error message if the code block fails
    print(paste("Error2 in block:", e))
  })
  }
}

# Print the results
head(forecast_df_girls)
print(m_2020_2050_girls)
```


# Regression of oafs

We now compute the average oao for 2020-2060 for each group. We can then use this to regress this average on the available average oaf of 2020-2050. This way, we can always plug in a number for obesity prevalence and  predict OAFs from that:

## requirements to fit oaf into projection dataframe:
```{r}
#country/countries defined above!
oaf_df_men<-data.frame(countries, Country, oafs)
oaf_df_men<-merge(oaf_df_men, m_2020_2050_men, by = "Country")

oaf_df_women<-data.frame(countries, Country, oafs)
oaf_df_women<-merge(oaf_df_women, m_2020_2050_women, by = "Country")

oaf_df_girls<-data.frame(countries, Country, oafs)
oaf_df_girls<-merge(oaf_df_girls, m_2020_2050_girls, by = "Country")

oaf_df_boys<-data.frame(countries, Country, oafs)
oaf_df_boys<-merge(oaf_df_boys, m_2020_2050_boys, by = "Country")

```
## regression for different groups:
Men:
```{r}
final<-lm(oaf_df_men$oafs~oaf_df_men$MeanProjection)
summary(final)
plot(oaf_df_men$oafs~oaf_df_men$MeanProjection)
abline(final$coefficients)
```

Women:
```{r}
final<-lm(oaf_df_women$oafs~oaf_df_women$MeanProjection)
summary(final)
plot(oaf_df_women$oafs~oaf_df_women$MeanProjection)
abline(final$coefficients)
```

Girls:
```{r}
final<-lm(oaf_df_girls$oafs~oaf_df_girls$MeanProjection)
summary(final)
plot(oaf_df_girls$oafs~oaf_df_girls$MeanProjection)
abline(final$coefficients)
```

Boys:
```{r}
final<-lm(oaf_df_boys$oafs~oaf_df_boys$MeanProjection)
summary(final)
plot(oaf_df_boys$oafs~oaf_df_boys$MeanProjection)
abline(final$coefficients)
```
## weighted lm to predict oafs:

We now compute the weighted average prevalence of every year and country. This way, we can fit one general model and not one specific one for each group. We use this final model to predict oafs.

##creating a dataframe for all groups:
```{r}
merged_df_adult <- merge(ag_population_adult, forecast_df_men, by.x=c("Location","Year"),by.y = c("Country", "Year"))
merged_df_adult <- merge(merged_df_adult, forecast_df_women, by.x=c("Location","Year"),by.y = c("Country", "Year"))

merged_df_child <- merge(ag_population_child, forecast_df_boys, by.x=c("Location","Year"),by.y = c("Country", "Year"))
merged_df_child <- merge(merged_df_child, forecast_df_girls, by.x=c("Location","Year"),by.y = c("Country", "Year"))

predictions_oao<-merge(merged_df_adult,merged_df_child, by=c("Location", "Year"))
predictions_oao$Population_girls <- predictions_oao$PopFemale.y
predictions_oao$Population_boys <- predictions_oao$PopMale.y
predictions_oao$PopMale.y <- NULL; predictions_oao$PopFemale.y<-NULL
predictions_oao$Population_men <- predictions_oao$PopMale.x
predictions_oao$Population_women <- predictions_oao$PopFemale.x
predictions_oao$PopMale.x<-NULL;predictions_oao$PopFemale.x<-NULL
```


Computing the weighted oao:
```{r}
predictions_oao$Forecast_weighted<-predictions_oao$Forecast_men*predictions_oao$Population_men/100+predictions_oao$Forecast_women*predictions_oao$Population_women/100+predictions_oao$Forecast_boys*predictions_oao$Population_boys/100+predictions_oao$Forecast_girls*predictions_oao$Population_girls/100
```

Dataframe for average oao and average oao to use for regression
```{r}
predictions_average_oaf_oao <- aggregate(Forecast_weighted ~ Location, data = predictions_oao, FUN = mean)
oaf_df_2060<-data.frame(countries, Country, oafs)
predictions_average_oaf_oao<-merge(oaf_df_2060, predictions_average_oaf_oao, by.x = "Country", by.y ="Location")
```

```{r}
final<-lm(predictions_average_oaf_oao$oafs~predictions_average_oaf_oao$Forecast_weighted)
summary(final)
plot(predictions_average_oaf_oao$oafs~predictions_average_oaf_oao$Forecast_weighted)
abline(final$coefficients)
```

# Direct medical costs

## Costs of 2016:
```{r}
#direct_medical_costs_2019<-read.csv("GHE_2019.csv")
#direct_medical_costs_2019<-direct_medical_costs_2019[-1,]
#colnames(direct_medical_costs_2019)[4] <- "GHE_2019"

#matrix<-cbind(1,predictions_oao$Forecast_weighted)
#predictions_average_oaf_oao<-matrix%*%final$coefficients
#average_weighted_oao_2060<-cbind(predictions_average_oaf_oao, average_weighted_oao_2060) #average_weighted_oao_2060 now includes the predicted oafs!
#direct_medical_costs_2019<-merge(direct_medical_costs_2019,average_weighted_oao_2060, by.x="Countries", by.y = "Location")
#direct_medical_costs_2019$GHE_2019 <- as.numeric(gsub(",", "", direct_medical_costs_2019$GHE_2019))
#direct_medical_costs_2019$DMC<-direct_medical_costs_2019$predictions_average_oaf_oao*direct_medical_costs_2019$GHE_2019
```
## costs till 2050/2060:

add oaf predictions to dataframe
```{r}
matrix<-cbind(1,predictions_oao$Forecast_weighted) #create matrix to predict oao
predictions_matrix_allyears_oaf<-matrix%*%final$coefficients #create predictions by using coefficients and matrix
predictions_oao<-cbind(predictions_matrix_allyears_oaf, predictions_oao)
#the predictions seem to make sense: compare them to oaf_df_2060
```

add forecasts of health spending and predict direct costs:
```{r}
direct_medical_costs_2050<-read.csv("IHME_EXPECTED_HEALTH_SPENDING_2019_2050_Y2021M09D22.csv") #only until 2050?
direct_medical_costs_2050<-direct_medical_costs_2050[, c(1,2,3,5,6)]
predictions_oao<-merge(direct_medical_costs_2050,predictions_oao, by.x=c("location_name", "year"), by.y = c("Location", "Year"))
predictions_oao$direct_costs_1000<-predictions_oao$predictions_matrix_allyears_oaf*predictions_oao$the_total_mean
```

#direct non-medical costs:

##travel cost:
We calculated this by multiplying the
average transport cost per trip by the population with overweight and obesity
```{r}
gas<-read.csv("e2373858-926a-4755-8350-78bb920f6ee9_Data.csv")
gas <- gas[, c(1,2,3,4,50,51,52,53,54,55,56,57,58, 59, 60, 61)]
gas$gas2016<-gas$X2016..YR2016.
#replace missing values by mean:
gas$gas2016[gas$gas2016 == ".."] <- NA
m <- mean(as.numeric(gas$gas2016), na.rm = TRUE)
gas$gas2016[is.na(gas$gas2016)] <- m
#reduce it to all countries
gas<-subset(gas, Country.Code %in% unique(data_oao$ISO))
```

```{r}
population<-read.csv("WPP2022_Demographic_Indicators_Medium.csv")
population<-population[, c(4,10,13,14,15,16,17)]
population<-population[population$Time>=2019, ]
#reduce it to all countries
population<-subset(population, ISO3_code %in% unique(data_oao$ISO))
#rename columns
colnames(population)[colnames(population) == "Time"] <- "Year"
colnames(population)[colnames(population) == "ISO3_code"] <- "ISO"
colnames(population)[colnames(population) == "TPopulation1Jan"] <- "Population_1000"

#reducing to 1975-2016:
#population<-population[population$Year >= "1975" & population$Year <="2016" ,]
#merge with data_oao
#merge_oao_population<-merge(data_oao, population, by = c("Year", "ISO"), all.x = TRUE)

```


We now add parameters for the number of outpatient and inpatient visits.
"parameter sourced from a scoping review of peer-reviewed studies!" Which peer-reviewed studies? "
"For outpatient visits andhospitalizations  (number and length), we use a global estimate due to the paucity of country-specific data" Which global estimate?

average inpatient/outpatient visits: (outpatient<-5.42, inpatient<-0.10)
)
https://www.researchgate.net/publication/329597721_Funding_and_services_needed_to_achieve_universal_health_coverage_applications_of_global_regional_and_national_estimates_of_utilisation_of_outpatient_visits_and_inpatient_admissions_from_1990_to_2016_a

HIC(spain):https://pubmed.ncbi.nlm.nih.gov/28261758/
HIC(UK): https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6180713/
HIC(US):https://journals.lww.com/md-journal/fulltext/2016/01120/body_mass_index_trajectories_and_healthcare.54.aspx
HIC(EU):https://www.dovepress.com/the-humanistic-and-economic-burden-associated-with-increasing-body-mas-peer-reviewed-fulltext-article-DMSO
HIC(Germany/italy):https://www.dovepress.com/obesity-in-germany-and-italy-prevalence-comorbidities-and-associations-peer-reviewed-fulltext-article-CEOR
LMIC(Brazil): https://dmsjournal.biomedcentral.com/articles/10.1186/s13098-018-0322-9


```{r}
##HIC:
#average outpatient visits per capita
#1.32=Frequency of GP visit per month, 1.30 = normal weight
GP<-1.32*12-1.30*12
Specialist<-1.38*12-1.3*12
hic_outpatient<-GP+Specialist
#average inpatient(hospitalisations) visits per capita
hospitalizations<-0.01
hic_inpatient<-hospitalizations
##LMIC:
#4.36=Frequency of GP visit per 6 months, 4.35 = normal weight
GP<-4.35*2-4.35*2
ER<-0.52*2-0.44*2
lmic_outpatient<-GP+ER
#average inpatient(hospitalisations) visits per capita
hospitalizations<-0.21*2-0.18*2
lmic_inpatient<-hospitalizations
#average outpatient visits per capita
```

We need the income groups if we want to use this data on inpatient/outpatient visits:
```{r}
income_groups<-read.csv("CLASS.csv")
predictions_oao<-merge(predictions_oao, income_groups[, c(2,4)], by.x="iso3", by.y="Code")
```
addinpatient/outpatient:
```{r}
predictions_oao$inpatient <- ifelse(predictions_oao$Income.group == "High income", hic_inpatient, lmic_inpatient)
predictions_oao$outpatient<- ifelse(predictions_oao$Income.group == "High income", hic_outpatient, lmic_outpatient)
```



### calculate for 2019-2060:
Before we do that We include it into our predictions_oao dataframe. We can then calculate gasprice * 5km(average distance to facility)*0.06(l/km). We can then calculate inpatient and outpatient costs.
```{r}
predictions_oao<-merge(predictions_oao, population[, c(2,3,4)], by.x=c("location_name", "year"), by.y=c("Location", "Year"))
predictions_oao$oaoPopulation_1000<-predictions_oao$Population_1000*predictions_oao$Forecast_weighted

predictions_oao<-merge(gas[, c(1,17)], predictions_oao, by.x="Country.Name", by.y="location_name", all.x=TRUE)

#gas price per l * km * l/km:
predictions_oao$atc<-as.numeric(predictions_oao$gas2016)*5*0.06
#inpatient/outpatient
outpatient_travel_cost<-predictions_oao$atc*predictions_oao$outpatient*predictions_oao$oaoPopulation
inpatient_travel_cost<-predictions_oao$atc*predictions_oao$inpatient*predictions_oao$oaoPopulation*2 #we multiply with two to also include informal caregiver travel costs!
predictions_oao$travel_cost_1000<-outpatient_travel_cost+inpatient_travel_cost
```
data clean up:
Our dataframe is getting messy so lets do a clean up:
```{r}
predictions_oao<-predictions_oao[, -c(4,8,9,11,12)]
```

#indirect costs:

##premature mortality like described:


This is calculated as the number
of years of potential life lost by individuals (by age group and sex) who died from OAO multiplied by the economic
value of a life year.

life expectancy/mortality rate:

##abriged version:

Data on life expectancy:
```{r}
life_expectancy_2100<-read.csv("WPP2022_Life_Table_Abridged_Medium_2022-2100.csv")

life_expectancy_2100<-life_expectancy_2100[, c(4, 13,16,17,19,20,10,28)]
life_expectancy_2022<-read.csv("WPP2022_Life_Table_Abridged_Medium_1950-2021.csv")
life_expectancy_2022<-life_expectancy_2022[life_expectancy_2022$Time >=2016, c(4,13,16,17,19,20,10,28)]
life_expectancy<-rbind(life_expectancy_2100, life_expectancy_2022)
life_expectancy <- life_expectancy[life_expectancy$Location %in% unique(data_oao$Country.Region.World), ]

life_expectancy$AgeGrp <- ifelse(life_expectancy$AgeGrp %in% c("95-99", "100+"), "95+", life_expectancy$AgeGrp) #this changes age group 95-99, 100+ to 95 as this is the way we have it in our ord dataframe below!
```

oao related deaths (ord):
```{r}
premature_mortality<-read.csv("ord_deaths_both.csv")
#premature_mortality_male<-read.csv("ord_deaths_male.csv") #read in first half of csv file
#premature_mortality_female<-read.csv("ord_deaths_female.csv") #second half
#premature_mortality<-rbind(premature_mortality_female, premature_mortality_male) #combine it

#We now need to add the relative risk(unfortunately manually due to no appropriate csv with the relative risks)
#We use the average relative risk of men and women due to scarcity of data!
premature_mortality$rr_values <- ifelse(premature_mortality$cause_name == "Ischemic stroke", 2.472,
                    ifelse(premature_mortality$cause_name == "Intracerebral hemorrhage", 3.066,
                           ifelse(premature_mortality$cause_name == "Subarachnoid hemorrhage", 3.066,
                                  ifelse(premature_mortality$cause_name == "Hypertensive heart disease", 3.122,
                                         ifelse(premature_mortality$cause_name == "Atrial fibrillation and flutter", 1.344,
                                             ifelse(premature_mortality$cause_name == "Asthma", 1.409,
                                                 ifelse(premature_mortality$cause_name == "Asthma", 1.409,
                                                        ifelse(premature_mortality$cause_name == "Diabetes mellitus type 2", 3.547,
                                                               ifelse(premature_mortality$cause_name == "Multiple myeloma", 1.09,
                                                                                                                                                    ifelse(premature_mortality$cause_name == "Leukemia", 1.111714286,
                                                                                                                                                           ifelse(premature_mortality$cause_name == "Alzheimer's disease and other dementias", 1.216,
                                                                                                                              ifelse(premature_mortality$cause_name == "Subarachnoid hemorrhage", 3.066,
                                                                                                                                     ifelse(premature_mortality$cause_name == "Hypertensive heart disease", 3.12,
                                                                                                                                            ifelse(premature_mortality$cause_name == "Ovarian", 1.038,
                                                                                                                                                   ifelse(premature_mortality$cause_name == "Kidney cancer", 1.28,
                                                        ifelse(premature_mortality$cause_name == "Thyroid cancer", 1.2585,1.998518  # Default value if no condition is met--> mean of RR of all other ord
                    )
)
)
)
)
)
)
)
)
)
)
)
)
)
)
)

#Attributable fraction:
premature_mortality$af<-(premature_mortality$rr_values-1)/premature_mortality$rr_values
#We can use this to calculate the ord deaths attributable to oao:
premature_mortality$deaths_oao <- premature_mortality$val*premature_mortality$af
```

change af to paf:

```{r}

oao_2019<-predictions_oao[predictions_oao$year==2019, c(1,17)]
#premature_mortality<-merge(oao_2019, premature_mortality, by.x="Country.Name", by.y="location_name" )
premature_mortality<-merge(premature_mortality, oao_2019, by.x="location_name", by.y="Country.Name" )

premature_mortality$af<-(premature_mortality$Forecast_weighted*(premature_mortality$rr_values-1))/(premature_mortality$Forecast_weighted*(premature_mortality$rr_values-1)+1)
premature_mortality$deaths_oao <- premature_mortality$val*premature_mortality$af


```



```{r}
ord_deaths<-premature_mortality[premature_mortality$measure_id==1, ]
ord_deaths<-premature_mortality[premature_mortality$metric_id==1, ]
#ord_deaths<-premature_mortality[premature_mortality$sex_name!="Both", ]
#ord_deaths<-premature_mortality[premature_mortality$sex_name=="Both", ]

ord_deaths$age_name <- gsub(" years", "", ord_deaths$age_name)
ord_deaths$age_name <- gsub(" year", "", ord_deaths$age_name)

ord_deaths$age_name[ord_deaths$age_name == "<1"] <- "0"
#ord_deaths$age_name[ord_deaths$age_name == "95+"] <- "95-99"

#sum up ords:
ord_deaths <- aggregate(cbind(val, deaths_oao) ~ sex_name + age_name + location_name + year, data = ord_deaths, FUN = sum)
#ord_deaths$sex_name <- ifelse(ord_deaths$sex_name=="Male", "Men", "Women")

#filter data on relevant countries/years:
ord_deaths <- ord_deaths[ord_deaths$location_name %in% unique(data_oao$Country.Region.World) & ord_deaths$year %in% 2016, ]
#life_expectancy <- life_expectancy[life_expectancy$Location %in% unique(data_oao$Country.Region.World), ]
country_iso<-unique(life_expectancy[, c(1,7)]) #add ISO codes to dataframe as some country-names may differ
ord_deaths<-merge(ord_deaths, country_iso, by.x="location_name", by.y="Location")
```

gdp per capita:
```{r}
gdp<-read.csv("GDP_worldbank.csv")
```

### Calculate costs with Given Life expetancy:

test it with Australia (with given life expectancy)
yll=life expectancy * deaths

```{r}
#ord_deaths_australia:
ord_deaths_australia<-ord_deaths[ord_deaths$location_name=="Australia", ]

life_expectancy_australia <- life_expectancy[life_expectancy$Location %in% "Australia" & life_expectancy$Sex %in% "Total" & life_expectancy$Time %in% 2016, ]
life_expectancy_australia[21,8]<-(life_expectancy_australia[21,8]+life_expectancy_australia[22,8])/2 #calculate average of last to as we have to remove age interval 100+
life_expectancy_australia<-life_expectancy_australia[-22, ] #remove last interval because we have 95+ twice
ord_deaths_australia$ylls<-life_expectancy_australia$ex*ord_deaths_australia$val
ord_deaths_australia$ylls_oao<-life_expectancy_australia$ex*ord_deaths_australia$deaths_oao

#try it:
sum_yl<-sum(ord_deaths_australia$ylls)
premature_cost_australia<-sum_yl*49876.7123761245 #mulitply ylls with value of life year
(sum_yl_australia_mult<-sum_yl)
(premature_australia_mult<-premature_cost_australia)

sum_yl_oao<-sum(ord_deaths_australia$ylls_oao)
premature_cost_australia_oao<-sum_yl_oao*49876.7123761245
sum_yl_oao
premature_cost_australia_oao
```
try it with given life expectation from all countries (loop through countries) with all deaths and not with only oao deaths (without RR):
```{r}
ord_deaths <- aggregate(val ~ age_name + location_name + year + ISO3_code, data = ord_deaths, FUN = sum)
life_expectancy <- life_expectancy[life_expectancy$Sex %in% "Total" & life_expectancy$Time %in% 2016, ]


result_df <- data.frame()
skip_countries <- c("TUR", "COK", "NIU", "TWN", "TKL")

for (country in unique(data_oao$ISO)) {
  if (country %in% skip_countries){
    next
  }
  print(country)
  ord_deaths_country <- ord_deaths[ord_deaths$ISO3_code == country, ]
  life_expectancy_country <- life_expectancy[life_expectancy$ISO3_code %in% country, ]
  life_expectancy_country <- life_expectancy_country[-22, ]  # remove last interval because we have 95+ twice

  gdp_country<-as.numeric(gdp[gdp$Country.Code==country, 48])

  # Calculate ylls for the country
  ylls <- life_expectancy_country$ex * ord_deaths_country$val

  # Add ylls to ord_deaths_country data frame
  ord_deaths_country$ylls <- ylls

  # Sum of ylls for the country
  sum_ylls_country <- sum(ylls)

  # Calculate premature cost for the country
  premature_cost_country <- sum_ylls_country * gdp_country  # multiply ylls with the value of a life year

  # Create a data frame with results for the country
  result_country <- data.frame(
    Location = country,
    Sum_YLLs = sum_ylls_country,
    Premature_Cost = premature_cost_country
  )

  # Append results to the main data frame
  result_df <- rbind(result_df, result_country)
}

# Print or use the result_df data frame containing results for each country
head(result_df)

```

### formula in PDF:

Let´s try it with the formula in like described (for australia:)
The results are a bit better but still much more...

test it with australia (with formula used)

withour RR:
```{r}
# Given data
total_yll <- 0

# Loop through each age group
h<-1
for (i in ord_deaths_australia$age_name) {
  ord_deaths_group <- ord_deaths_australia[ord_deaths_australia$age_name == i, ]
  c<-life_expectancy_australia$mx[h:length(life_expectancy_australia$mx)]
  
  #this is basically the formula:
  a <- ord_deaths_group$val[1] * cumprod(1- c)
  b<-sum(a)+sum(a[length(a)]*(1-life_expectancy_australia$mx[21]^100))
  # Add YPLL to the total
  total_yll <- total_yll + b
  
  # Print results for each age group
  cat("AgeGroup", i, "\n")
  cat("cumprod", c, "\n")
  cat("b", b, "\n")
  h<-h+1
}

# Print total YPLL
cat("Total YLL:", total_yll, "\n")
cost<-total_yll*49876.7123761245
cat("Total cost australia:", cost, "\n")

sum_yl_australia_form<-total_yll
premature_australia_form<-cost
```
with RR:

```{r}
# Given data
total_yll <- 0

# Loop through each age group
h<-1
for (i in ord_deaths_australia$age_name) {
  ord_deaths_group <- ord_deaths_australia[ord_deaths_australia$age_name == i, ]
  c<-life_expectancy_australia$mx[h:length(life_expectancy_australia$mx)]
  
  #this is basically the formula:
  a <- ord_deaths_group$deaths_oao[1] * cumprod(1- c)
  b<-sum(a)+sum(a[length(a)]*(1-life_expectancy_australia$mx[21]^100))
  # Add YPLL to the total
  total_yll <- total_yll + b
  
  # Print results for each age group
  cat("AgeGroup", i, "\n")
  cat("cumprod", c, "\n")
  cat("b", b, "\n")
  h<-h+1
}

# Print total YPLL
cat("Total YLL:", total_yll, "\n")
cost<-total_yll*49876.7123761245
cat("Total cost australia:", cost, "\n")

sum_yl_australia_form_oao<-total_yll
premature_australia_form_oao<-cost
```



##complete version:
```{r}
#life expectancy:
life_expectancy_2100<-read.csv("WPP2022_Life_Table_Complete_Medium_Both_2022-2100.csv")

life_expectancy_2100<-life_expectancy_2100[, c(13,16,17,19,20,10,28)]
life_expectancy_2022<-read.csv("WPP2022_Life_Table_Complete_Medium_Both_1950-2021.csv")
life_expectancy_2022<-life_expectancy_2022[life_expectancy_2022$Time >=2016, c(13,16,17,19,20,10,28)]
life_expectancy<-rbind(life_expectancy_2100, life_expectancy_2022)
```
oao related deaths (ord):
```{r}
premature_mortality<-read.csv("ord_deaths_both.csv")
#premature_mortality_male<-read.csv("ord_deaths_male.csv") #read in first half of csv file
#premature_mortality_female<-read.csv("ord_deaths_female.csv") #second half
#premature_mortality<-rbind(premature_mortality_female, premature_mortality_male) #combine it
ord_deaths<-premature_mortality[premature_mortality$measure_id==1, ]
ord_deaths<-premature_mortality[premature_mortality$metric_id==1, ]
#ord_deaths<-premature_mortality[premature_mortality$sex_name!="Both", ]
ord_deaths$age_name <- gsub(" years", "", ord_deaths$age_name)
ord_deaths$age_name[ord_deaths$age_name == "<1"] <- "0"

#sum up ords:
ord_deaths <- aggregate(val ~ sex_name + age_name + location_name + year, data = ord_deaths, FUN = sum)
ord_deaths$sex_name <- ifelse(ord_deaths$sex_name=="Male", "Men", "Women")

#filter data on relevant countries/years:

#ord_oao<-merge(ord_deaths, data_oao, by.x=c("year", "location_name", "sex_name"), by.y=c("Year", "Country.Region.World","Sex"))
ord_deaths <- ord_deaths[ord_deaths$location_name %in% unique(data_oao$Country.Region.World) & ord_deaths$year %in% 2016, ]
life_expectancy <- life_expectancy[life_expectancy$Location %in% unique(data_oao$Country.Region.World), ]
```
test it with australia male:
```{r}
#ord_deaths_australia:
ord_deaths_australia<-ord_deaths[ord_deaths$location_name=="Australia", ]
ord_deaths_australia_male<-ord_deaths[ord_deaths$location_name=="Australia"&ord_deaths$sex_name=="Men", ]

life_expectancy_australia_male <- life_expectancy[life_expectancy$Location %in% "Australia" & life_expectancy$Sex %in% "Total" & life_expectancy$Time %in% 2016, ]
life_expectancy_australia_male<-life_expectancy_australia_male[-22, ] #remove last interval becaue we have 95+ twice
ord_deaths_australia_male$ylls<-life_expectancy_australia_male$ex*ord_deaths_australia_male$val
ylls<-ord_deaths_australia_male$val*life_expectancy_australia_male$ex

#try it:
sum_yl<-sum(ord_deaths_australia_male$ylls)
premature_cost_australia<-sum_yl*520*4
#try it with the formula used in pdf:
3.09*cumprod(1-life_expectancy_australia_male$mx)
sum(3.09*cumprod(1-life_expectancy_australia_male$mx))*4
```





##premature mortality with calculated Ylls:

We use calculated YLLs from GDP 2019 for calculations. Due to missing data on RR we just projected the value form 2019 by multiplying it with the relative change of oao for the projected years. (?)

YLLs:

```{r}
ylls_australia<-read.csv("IHME-GBD_2019_DATA-16794036-1.csv")
ylls_australia<-ylls_australia[ylls_australia$metric_id==1, ]
ylls_australia<-ylls_australia[ylls_australia$measure_id==4, ]
ylls_australia<-ylls_australia[ylls_australia$sex_name=="Both", ]
ylls_australia$rr_values <- ifelse(ylls_australia$cause_name == "Ischemic stroke", 2.472,
                    ifelse(ylls_australia$cause_name == "Intracerebral hemorrhage", 3.066,
                           ifelse(ylls_australia$cause_name == "Subarachnoid hemorrhage", 3.066,
                                  ifelse(ylls_australia$cause_name == "Hypertensive heart disease", 3.122,
                                         ifelse(ylls_australia$cause_name == "Atrial fibrillation and flutter", 1.344,
                                             ifelse(ylls_australia$cause_name == "Asthma", 1.409,
                                                 ifelse(ylls_australia$cause_name == "Asthma", 1.409,
                                                        ifelse(ylls_australia$cause_name == "Diabetes mellitus type 2", 3.547,
                                                              ifelse(ylls_australia$cause_name == "Multiple myeloma", 1.09,
                                                                                                                                                    ifelse(ylls_australia$cause_name == "Leukemia", 1.111714286,
                                                                                                                                                           ifelse(ylls_australia$cause_name == "Alzheimer's disease and other dementias", 1.216,
                                                                                                                              ifelse(ylls_australia$cause_name == "Subarachnoid hemorrhage", 3.066,
                                                                                                                                     ifelse(ylls_australia$cause_name == "Hypertensive heart disease", 3.12,
                                                                                                                                            ifelse(ylls_australia$cause_name == "Ovarian", 1.038,
                                                                                                                                                   ifelse(ylls_australia$cause_name == "Kidney cancer", 1.28,
                                                        ifelse(ylls_australia$cause_name == "Thyroid cancer", 1.2585,1.998518  # Default value if no condition is met--> mean of RR of all other ord
                    )
)
)
)
)
)
)
)
)
)
)
)
)
)
)
)
```

```{r}
oao_2019<-predictions_oao[predictions_oao$year==2019, c(1,17)]
ylls_australia<-merge(ylls_australia, oao_2019, by.x="location_name", by.y="Country.Name" )

ylls_australia$af<-(ylls_australia$Forecast_weighted*(ylls_australia$rr_values-1))/(ylls_australia$Forecast_weighted*(ylls_australia$rr_values-1)+1)
ylls_australia$oao_yll <- ylls_australia$val*ylls_australia$af
(sum_yl_australia_oao_gbd<-sum(ylls_australia$oao_yll))
(premature_australia_oao_gbd<-sum(ylls_australia$oao_yll)*49876.7123761245)
(sum_yl_australia_gbd<-sum(ylls_australia$val))
(premature_australia_gbd<-sum(ylls_australia$val)*49876.7123761245)

```

Let´s compare the different methods:
```{r}
cat("Ylls when using life expectancy * deaths:", sum_yl_australia_mult/10^9,"\n")
cat("Costs when using life expectancy * deaths:", premature_australia_mult/10^9, "\n")
cat("Ylls when using life expectancy * deaths with RR:", sum_yl_oao/10^9,"\n")
cat("Costs when using life expectancy * deaths with RR:", premature_cost_australia_oao/10^9, "\n")

cat("Ylls when using formula from authors:", sum_yl_australia_form/10^9, "\n")
cat("Costs when using formula from authors:",premature_australia_form/10^9, "\n")
cat("Ylls when using formula from authors with RR:", sum_yl_australia_form_oao/10^9, "\n")
cat("Costs when using formula from authors with RR:",premature_australia_form_oao/10^9, "\n")

cat("Ylls when using Ylls from GBD:", sum_yl_australia_gbd/10^9,"\n")
cat("Costs when using Ylls from GBD:", premature_australia_gbd/10^9,"\n")
cat("Ylls when using Ylls from GBD with RR:", sum_yl_australia_oao_gbd/10^9,"\n")
cat("Costs when using Ylls from GBD with RR:", premature_australia_oao_gbd/10^9,"\n")
```
men,women,boy,girls-wise:?
```{r}
#Using aggregate to sum values by sex and age_group
ylls$sex_name <- ifelse(ylls$sex_name=="Male", "Men", "Women")
ylls <- aggregate(val ~ sex_name + age_name + location_name + year, data = ylls, FUN = sum)

#for age group <20:
ylls_child<-ylls[ylls$age_name=="<20 years",]
ylls_child$sex_name <- ifelse(ylls_child$sex_name=="Men", "Boys", "Girls")
#combine it with data_oao_child:
data_oao_child<-merge(ylls_child, data_oao_child, by.x=c("year", "location_name", "sex_name"), by.y=c("Year", "Country", "Sex"))

#for age group >20:
ylls_adult<-ylls[ylls$age_name=="20+ years",]
#combine it with data_oao_adult:
data_oao_adult<-merge(ylls_adult, data_oao, by.x=c("year", "location_name", "sex_name"), by.y=c("Year", "Country.Region.World", "Sex"))
```
```{r}
#for age group <20:
data_oao_child<-merge(gdp[, c(1, 48)], data_oao_child, by.x=c("Country.Name"), by.y=c("location_name"))
data_oao_child$X2016..YR2016.<-as.numeric(data_oao_child$X2016..YR2016.)
data_oao_child$premature_mortality<-data_oao_child$X2016..YR2016.*data_oao_child$val

#for age group >20:
data_oao_adult<-merge(gdp[, c(1, 48)], data_oao_adult, by.x=c("Country.Name"), by.y=c("location_name"))
data_oao_adult$X2016..YR2016.<-as.numeric(data_oao_adult$X2016..YR2016.)
data_oao_adult$premature_mortality<-data_oao_adult$X2016..YR2016.*data_oao_adult$val

```



##reduced productivity due to additional absenteeism:


```{r}
income<-read.csv("Earnings2019.csv")
income<-income[income$classif1.label=="Economic activity (Aggregate): Total", ]
income<-income[income$classif2.label=="Currency: U.S. dollars", ]
income<-income[income$ref_area.label %in% unique(data_oao$Country.Region.World), ]


income<-read.csv("cf68b894-02db-4037-b35e-8b66af165f6b_Data.csv")
income<-income[income$Country.Name %in% unique(data_oao$Country.Region.World), ]
income[206, 5]<-"36072.7" #not viewable in View? Add value from United kingdom of last year!
income$daily_2019<-as.numeric(income$X2019..YR2019.)/248 #divide through average number of workdays
```
employment rates:
```{r}
employment<-read.csv("employment2019.csv")

#use data from previous years if NA:
employment$employment_rate<-ifelse(employment$X2019..YR2019.=="..", employment$X2018..YR2018., employment$X2019..YR2019.)

employment$employment_rate<-ifelse(employment$employment_rate=="..", employment$X2017..YR2017., employment$employment_rate)

employment<-employment[employment$Country.Name %in% unique(data_oao$Country.Region.World), ]
employment$employment_rate<-as.numeric(employment$employment_rate)/100
```
add abstenteeism:

"parameter sourced from a scoping review of peer-reviewed studies! https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4225558
https://journals.lww.com/joem/abstract/2014/01000/sick_leave_days_and_costs_associated_with.3.aspx

LIC: https://pubmed.ncbi.nlm.nih.gov/29568332/
Specifcally, for the normal BMI group the adjusted means for absenteeism was 6.21% and it was signifcantly lower than the overweight group (6.46; p=0.02), obesity class I (6.82; p<0.001), obesity class II (8.28; p<0.001), and obesity class III (10.03;
p<0.001)

create weighted average:
Among all respondents (n=34,254), a total of 46.4% were
normal weight, 34.9% were overweight, and the remaining 18.7% were obese (12.9% were obese class I, 3.7% were obese class II, and 2.1% were obese class III)"

HIC:https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0233512#pone-0233512-t001

LIC:
```{r}
#parameter sourced from a scoping review of peer-reviewed studies! https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4225558
#https://journals.lww.com/joem/abstract/2014/01000/sick_leave_days_and_costs_associated_with.3.aspx
total<-53.6
obesity_class_I<-12.9/total
obesity_class_II<-3.7/total
obesity_class_III<-2.1/total
overweight<-34.9/total

prevalence_rates<-c(overweight, obesity_class_I,obesity_class_II, obesity_class_III)
absenteeism_rates <- c(6.46, 6.82, 8.28, 10.03)


# Calculate the weighted average
weighted_average <- sum(absenteeism_rates * prevalence_rates) / sum(prevalence_rates)

lic_additional_days_absent<-((weighted_average-6.21)*248)/100 #substract average absenteiism and multiple with average working days to get additional days absent
lic_additional_days_absent
```

HIC:
```{r}
#calculate relative values of all classes for weighted average
total<-117.769
obesity_class_I<-14.749/total
obesity_class_II<-5.052/total
obesity_class_III<-2.334/total
overweight<-34.532/total

prevalence_rates<-c(overweight, obesity_class_I,obesity_class_II, obesity_class_III)
absenteeism_rates<-c(0.71, 1.14, 1.23, 1.79)

weighted_average<-sum(absenteeism_rates*prevalence_rates)/sum(prevalence_rates)
hic_additional_days_absent<-(weighted_average-0.5)*248/100
hic_additional_days_absent
```

add these to the dataframe:
```{r}
employment_income<-merge(income, employment[, c(1,2,8)], by="Country.Name")
#average_daily_wage<-result_income$daily_2016
data_oao_adult<-rbind(data_oao_men, data_oao_women)
data_oao_adult<-data_oao_adult[data_oao_adult$Year=="2016", ]
data_oao_adult<-aggregate(oao~Country.Region.World+ISO, data = data_oao_adult, mean) #take weighted mean instead of mean!

employment_income_oao<-merge(employment_income, data_oao_adult, by.x="Country.Name", by.y="Country.Region.World")
employment_income_predictions<-merge(employment_income,predictions_oao, by="Country.Name")

#lic_additional days
employment_income_predictions$absenteeism_cost2<-(employment_income_predictions$oaoPopulation_1000*1000)*employment_income_predictions$employment_rate*lic_additional_days_absent*employment_income_predictions$daily_2019
absenteeism_cost_lic<-(employment_income_predictions$oaoPopulation_1000*1000)*employment_income_predictions$employment_rate*lic_additional_days_absent*employment_income_predictions$daily_2019
#hic_additional days
absenteeism_cost_hic<-(employment_income_predictions$oaoPopulation_1000*1000)*employment_income_predictions$employment_rate*hic_additional_days_absent*employment_income_predictions$daily_2019

#add it based on income level
employment_income_predictions$absenteeism<-ifelse(employment_income_predictions$Income.group == "High income", hic_additional_days_absent, lic_additional_days_absent)
employment_income_predictions$absenteeism_cost<-employment_income_predictions$oaoPopulation_1000*1000*employment_income_predictions$employment_rate*employment_income_predictions$absenteeism*employment_income_predictions$daily_2019
```


rename columns:
```{r}
colnames(predictions_oao)[colnames(predictions_oao) == "direct_costs"] <- "direct_costs(thousands)"
colnames(predictions_oao)[colnames(predictions_oao) == "the_total_mean"] <- "Total_health_expenditure_1000"
colnames(predictions_oao)[colnames(predictions_oao) == "travel_cost"] <- "travel_cost(thousands)"
colnames(population)[colnames(population) == "TPopulation1Jan"] <- "Population_1000"
colnames(predictions_oao)[colnames(predictions_oao) == "oaoPopulation"] <- "oaoPopulation(thousands)"
```


Wage predictions missing!

##reduced productivity due to additional presenteeism:


